---
title: "p8105_hw3_ss6692"
output: github_document
date: "2023-10-11"
---
## Problem 2
```{r libraries, message=FALSE}
library(tidyverse)
library(readxl)
library(dplyr)
library(janitor)
library(tidyr)
library(knitr)
library(viridis)
```
## Problem 2
### Loading the dataset
```{r load_brfss_smart2010}
library(p8105.datasets)
data("brfss_smart2010")
```
### Here, we are cleaning the data and filtering for "Overall Health." We will then order the responses from Poor to Excellent.
```{r clean_and_order_brfss_df}
brfss_df = brfss_smart2010 |>
  janitor::clean_names()|>
  select(-locationabbr)|>
  separate(locationdesc, into = c("state", "county"), sep = " - ")|>
    filter(topic == "Overall Health" & response %in% c("Excellent", "Very good", "Good", "Fair", "Poor")) |>
    mutate(response = factor(response, levels = c("Poor", "Fair", "Good", "Very good", "Excellent"), ordered = TRUE))
```
## Now, we will find which states which states were observed at 7 or more locations in 2002.
```{r states_2002}
brfss_2002_df = 
   brfss_df|>
    filter(year == 2002) |>
    group_by(state) |>
    summarize(count = n()) |>
    filter(count >= 7)
kable(brfss_2002_df)
```
## Now we are finding the same information for 2010.
```{r states_2010}
brfss_2010_df = 
   brfss_df|>
    filter(year == 2010) |>
    group_by(state) |>
    summarize(count = n()) |>
    filter(count >= 7)
kable(brfss_2010_df)
```
In 2002, the following states ` r kable(brfss_2002_df)` were observed at 7 or more locations.

### Now, we are creating a dataset containing only "Excellent" responses. We will also add a new variable that averages `data_value` across locations, but within states.

```{r excellent_df}
excellent_df =
  brfss_df |>
  filter(response == "Excellent")|>
group_by(state, year) |>
  summarize(avg_data_value = mean(data_value, na.rm = TRUE))
```
### We will use the excellent_df to create a spaghetti plot
```{r excellent_plot}
excellent_df |> 
  ggplot(aes(x = year, y = avg_data_value, color = state)) + 
  geom_line() +
  labs(title = "Mean data values of Excellent responses by state ", 
        x = "Year",
        y = "Mean Data Value")
```

```{r states_plot}
nyc_df =
  brfss_df |>
  filter((year == 2006 | year == 2010) & state == "NY") |>
  select(year, response, data_value, county)
  
nyc_df |>
  ggplot(aes(x=response, y=data_value)) +
  geom_point() +
  facet_grid(. ~ year) + 
  labs(title = "NY state health responses in 2006 and 2010", 
        x = "Response",
        y = "Data Value") 
```

## Problem 3
```{r load_accel_df}
nhanes_accel_df = read_csv(file = "data/nhanes_accel.csv")|>
  janitor::clean_names()|>
 pivot_longer(
    min1:min1440,
    names_to = "minute",
    names_prefix = "min",
    values_to = "mims"
  )
  
```

```{r load_covar_df}
nhanes_covar_df = read_csv(file = "data/nhanes_covar.csv",skip = 4, na = "." )|>
  janitor::clean_names() |>
 mutate(
    sex = ifelse(sex == 1, "male", "female"), 
    education = case_match(education,
                           "1" ~ "less than high school", 
                           "2" ~ "high school equivalent", 
                           "3" ~ "more than high school"), 
    education = fct_relevel(education, c("less than high school", "high school equivalent", "more than high school"))
    )|> 
      filter(age >= 21) |> 
      drop_na()
```

```{r merge_accel_covar}
MIMS_df = 
  inner_join(nhanes_covar_df, nhanes_accel_df)
```
### Now we are creating a table to show the number of males and females in each educatio group.
```{r education_table_by_sex}
education_df = MIMS_df |>
  group_by(education, sex) |>
  summarize(count = n()) |>
  pivot_wider(names_from = sex, values_from = count)
```
### Now, we will create a visualization of the age distributions for men and women in each education category.
```{r education_boxplot}
ggplot(MIMS_df, aes(x = education, y = age, fill = sex)) +
  geom_boxplot() +
  labs(title = "Age distributions by education and sex",
       x = "education",
       y = "age") +
  theme_minimal()
```
As shown by the box plot, males and females with more than a high school education have the lowest median age, and females with less than a high school education and high school equivalent education have the highest median age. The age distribution of both males and females with more than a high school education is right skewed, and the distribution of both males and females of a high school equivalent education, and females with less than a high school education is slightly left skewed. The age distribution of males with less than a high school education is fairly normally distributed.
### Now, we will make a plot to show the distributin of total activity by sex and education level.
```{r}
MIMS_df |> 
  group_by(seqn, age, sex, education) |> 
  summarize(total_activity = sum(mims)) |> 
  ggplot(aes(x = age, y = total_activity, color = sex)) + 
  geom_point() +
  geom_smooth() +
  facet_grid(. ~ education) + 
  labs(title = "Total activity by sex and education level", 
        x = "Age",
        y = "Total Activity (MIMS)") +
    scale_color_viridis_d() 
```

### Finally, we are making a plot to visualize 24-hr activity time courses by education and sex.
```{r}
MIMS_df |>
  mutate(
    minute = as.numeric(minute)
  ) |> 
ggplot(aes(x = minute, y = mims, color = sex)) +
   geom_point(alpha = 0.5) +
  facet_grid(. ~ education) + 
  labs(title = "24-hr activity time courses by education and sex", 
        x = "Minute of the Day",
        y = "Activity (mims)") +
      scale_color_viridis_d() 
```




